---
title: "Driver_Performance_Evaluation"
author: "Micah Turner"
date: "2023-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Scope

The current (April, 2023) monthly driver evaluation process features export of a csv file from a relational database containing a summary of driver metrics over a specified time period (typically one calender month). Individual driver performance metrics are manually extracted from this summary and compiled into a report featuring the following information for each day of the specified time period, in addition to aggregates as depicted below;
- Completed moves (cms)
- Completed moves aggregate 
- Moves per hour (mph)
- Moves per hour aggregate 
- On time percent (otp)
- On time percent aggregate 

The purpose of this project is to improve this report by adding shift averages for the above metrics, adding hours worked (to track missed days, early departures, and extra shifts), and then automate the production of this report using the R programming langauge. This project is on a volunteer basis without financial compensation / allocated billable hours, for the purpose of gaining experience and going above and beyond.


# Approach

While I expect there is a more efficient method, this is my first independent project using R and I am breaking the problem down into it's individual components and then solving them one by one. The approach features four steps;
- Step 1: Load the libraries required for the solution
- Step 2: Read the csv file into R and prepare it for processing
- Step 3: Create a custom function (driver_eval) that takes a driver ID as an argument and outputs a ready to print pdf file containing the report as defined by the project scope
- Step 4: Create a "for loop" that takes a vector containing all the driver IDs and runs them through the driver_eval function

All four steps are combined into a single "code chunk" for the purpose of minimizing the number of manual inputs required to produce the reports, however, they were created piece by piece (see GitHub version history). The appendix contains partition "code chunks" to make debugging easier. 

Also of note, calender months have a variable number of days while the number of drivers fluctuates over time, and a design goal is to turn this process over to operational users who do not understand the code. Therefore, a derived requirement for the code is to work regardless of the number of days in the summary and the number of drivers contained in that summary so the end user is not required to manually adjust either variable. 

# Code
# Load Libraries
```{r}
## Step 1
library(tidyr)
library(data.table)
library(dplyr)
library(gridExtra)
```

```{r}

## Step 2: Read the csv file into R and prepare it for processing

### Import Summary
gid <- "1yh--wWbGDo0cfX-CzSV9Qrlb-9r76JGv" #Google Drive document ID
smry_v1 <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", gid), header=FALSE)

### Create a_number vector: Driver ID's are referred to as "A numbers" because they start with that letter
a_number_v1 <- pull(smry_v1,V2)
a_number_v2 <- unique(a_number_v1)
a_number <- a_number_v2[3:length(a_number_v2)]

### Remove rows below a_number in smry_v1 dataframe
c1 <- length(a_number)+3
c2 <- nrow(smry_v1)
smry_v2 <- smry_v1[-c(c1:c2),]

### Remove commas, which if present returns NA upon conversion to numeric class (e.g 1,215 to 1215)
c3 <- c(3:length(smry_v2))
smry_v2[c3] <- lapply(smry_v2[c3], gsub, pattern =",", replacement="")

### Remove % symbols, which if present returns NA upon conversion to numeric class (e.g. 98% = 98)
smry_v2[c3] <- lapply(smry_v2[c3], gsub, pattern ="%", replacement="")

### Completed Moves
cms_v1 <- smry_v2[, smry_v2[2, ] == "Completed Moves"]
cms_v2 <- smry_v2[, c(2,length(smry_v2)-3)]
cms_v3 <- cbind(cms_v2, cms_v1)

### Moves per Hour
mph_v1 <- smry_v2[, smry_v2[2, ] == "MPH"]
mph_v2 <- smry_v2[, c(2,length(smry_v2)-1)]
mph_v3 <- cbind(mph_v2, mph_v1)

### On Time Percentage
otp_v1 <- smry_v2[, smry_v2[2, ] == "On-Time Moves "]
otp_v2 <- smry_v2[, c(2,length(smry_v2))]
otp_v3 <- cbind(otp_v2, otp_v1)

### Hours Worked
hrw_v1 <- smry_v2[, smry_v2[2, ] == "Hours Worked"]
hrw_v2 <- smry_v2[, c(2,length(smry_v2)-2)]
hrw_v3 <- cbind(hrw_v2, hrw_v1)

### Make row 1, which contains dates, the header row and then delete row 1
cms_v4 <- cms_v3
colnames(cms_v4) <- cms_v3[1, ]
cms_v4 = cms_v4[-1,]

mph_v4 <- mph_v3
colnames(mph_v4) <- mph_v3[1, ]
mph_v4 = mph_v4[-1,]

otp_v4 <- otp_v3
colnames(otp_v4) <- otp_v3[1, ]
otp_v4 = otp_v4[-1,]

hrw_v4 <- hrw_v3
colnames(hrw_v4) <- hrw_v3[1, ]
hrw_v4 = hrw_v4[-1,]

### Label the column containing A numbers as such
colnames(cms_v4)[1] <- "Driver_ID"
colnames(mph_v4)[1] <- "Driver_ID"
colnames(otp_v4)[1] <- "Driver_ID"
colnames(hrw_v4)[1] <- "Driver_ID"

### Label the aggregate column appropriately
colnames(cms_v4)[2] <- "Total"
colnames(mph_v4)[2] <- "Total"
colnames(otp_v4)[2] <- "Total"
colnames(hrw_v4)[2] <- "Total"

### Change the data class from character to numeric
c4 <- c(2:length(cms_v4))
cms_v4[ ,c4] <- apply(cms_v4[ ,c4], 2, function(x) as.numeric(as.character(x)))
mph_v4[ ,c4] <- apply(mph_v4[ ,c4], 2, function(x) as.numeric(as.character(x)))
otp_v4[ ,c4] <- apply(otp_v4[ ,c4], 2, function(x) as.numeric(as.character(x)))
hrw_v4[ ,c4] <- apply(hrw_v4[ ,c4], 2, function(x) as.numeric(as.character(x)))

### Create new dataframe
cms_v5 <- cms_v4
mph_v5 <- mph_v4
otp_v5 <- otp_v4

### Calculate mean of all numeric columns
cms_v5 <- cms_v4 %>% summarize_if(is.numeric, mean, na.rm = TRUE)
mph_v5 <- mph_v4 %>% summarize_if(is.numeric, mean, na.rm = TRUE)
otp_v5 <- otp_v4 %>% summarize_if(is.numeric, mean, na.rm = TRUE)

### Label the new row 
Driver_ID <- "Average"
cms_v5$Driver_ID <- Driver_ID
mph_v5$Driver_ID <- Driver_ID
otp_v5$Driver_ID <- Driver_ID

### Reorder A number column from last to first
cms_v5 <- cms_v5[, c(length(cms_v5),1:length(cms_v5)-1)]
mph_v5 <- mph_v5[, c(length(mph_v5),1:length(mph_v5)-1)]
otp_v5 <- otp_v5[, c(length(otp_v5),1:length(otp_v5)-1)]

### In the darkness bind them
cms_v6 <- rbind(cms_v4,cms_v5)
mph_v6 <- rbind(mph_v4,mph_v5)
otp_v6 <- rbind(otp_v4,otp_v5)

### Round to appropriate digit
cms_v7 <- cms_v6 %>% mutate(across(is.numeric, round, digits=1))
mph_v7 <- mph_v6 %>% mutate(across(is.numeric, round, digits=1))
otp_v7 <- otp_v6 %>% mutate(across(is.numeric, round, digits=1))

### Clean up variable names
cms <- cms_v7
mph <- mph_v7
otp <- otp_v7
hrw <- hrw_v4

## Step 3

driver_eval <- function(anumber){
             ### Extract each metric
        acms <- cms[cms$Driver_ID %in% c(anumber,"Average"),]
        amph <- mph[mph$Driver_ID %in% c(anumber,"Average"),]
        aotp <- otp[otp$Driver_ID %in% c(anumber,"Average"),]
        ahrw <- hrw[hrw$Driver_ID==anumber,]
        
        ### Create vector for header row
        id <- c("Completed Mvs", "Avg Completed Mvs", "MPH", "Avg MPH", "On Time %", "Avg On Time %", "Hrs Worked")
        
        ### In the darkness bind them
        ar1 <- rbind(acms, amph, aotp, ahrw)
        ar2 <- cbind(id, ar1)
        ar2 = ar2[,-2]
        
        ### Pivot long and wide
        p1 <- colnames(ar2)
        p2 <- p1[!p1 == "id"]
        ar3 <- ar2 %>%
                 pivot_longer(
                         cols = p2,
                         names_to = "date",
                         values_to = "value"
      
        )
        ar4 <- ar3 %>%
                pivot_wider(
                        names_from = id,
                        values_from = value
        )
        
        ### Cleanup
        colnames(ar4)[1] <- anumber
        ar5 <- na.omit(ar4)
        
        ### Export to PDF
        pdf(paste(Sys.Date(),"_",anumber,".pdf"), height = 8.5, width = 11)
        grid.table(ar5) 
        dev.off()
}

## Step 4: For Loop
for(i in a_number){
driver_eval(i)
}   

```

```{r}
mph_barchart <- ggplot(smry_v5, aes(x=a_number,y=mph_tgt2)+
        geom_bar(stat = "identity")
mph_barchart
```