---
title: "The Absent Opportunity"
author: "Micah Turner"
date: "2023-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Absent Opportunity
This analysis explores the concept of using unplanned absences from work derived from a variety of factors like the retention rate and sick days as an opportunity to increase productivity leveraging data driven decision making.

## Background
There are two productivity metrics used to evaluate driver performance. The first of these is moves per hour (MPH), where the goal is to achieve 5 mph. The second driver performance metric is On Time Percent (OTP), where the goal is to achieve 78% of moves on time. 

```{r}
### Feb, 2023 MPH
smry_v3 <- smry_v3 %>%
  mutate(ID = fct_reorder(ID, MPH))

mph_barchart <- ggplot(smry_v3) +
  geom_bar(aes(x = ID, y = MPH),stat = "identity") +
  geom_hline(yintercept = 12.2, color = "red") +
  coord_flip()+
  annotate("text", x = 5, y = 8, label = "4% of drivers meet performance standard")+
  labs(title = "Feb, 2023 MPH")

mph_barchart
```


```{r}
### Feb, 2023 MPH
smry_v3 <- smry_v3 %>%
  mutate(ID = fct_reorder(ID, Tgt_Moves))

target_barchart <- ggplot(smry_v3, aes(x = ID, y = Tgt_Moves, fill=Qty_OnTime)) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip()+
  scale_fill_gradient2(low = "black", high = "red")+
  labs(title = "Feb, 2023 Performance Gap")

target_barchart
```

```{r}
s2
```

## Data


```{r}
### Extract the 2nd column (total) from the dataframes constructed in the driver_eval function and make them the same length
mph_tgt <- mph[,2]
mph_tgt2 <- mph_tgt[2:28]
otp_tgt <- otp[,2]
otp_tgt2 <- otp_tgt[2:28]
cms_tgt <- cms[,2]
cms_tgt2 <- cms_tgt[2:28]
hrw_tgt <- hrw[,2]
hrw_tgt2 <- hrw_tgt[2:28]

### Create a vector with calculation results for the number of moves required to bring a driver's mph equal to 5
target <- sapply(1:length(a_number), function(x) solve(1/hrw_tgt2[x], 5-cms_tgt2[x]/hrw_tgt2[x]))

### Create an intermediary vector to be used for a follow on calculation - this calculates the total number of on time moves
otm <- cms_tgt2*otp_tgt2/100

### Get rid of decimal in the target vector
target1 <- round(target, digits = 0)

### Calculate the number of target moves that need to be on time to stay within the 78% on time performance metric
s2 <- sapply(1:length(a_number), function(x) {
  solve(1, .78 * (cms_tgt2[x] + target[x]) - otm[x])
})

### Get rid of the decimal in the s2 vector
s2 <- round(s2, digits = 0)

### Replace negative values with 0
target1 <- pmax(target1,0)
s2 <- pmax(s2,0)

### Trim empty spaces from vectors
target1 <- trimws(target1)
s2 <- trimws(s2)

### In the darkness bind them
smry_v3 <- cbind(a_number,mph_tgt2,target1,s2)

### Rename Columns
new_col_names <- c("ID","MPH","Tgt_Moves","Qty_OnTime")
colnames(smry_v3) <- new_col_names

### Change type to dataframe
smry_v3 <- as.data.frame(smry_v3)

###Make them numeric again
smry_v3[ ,-1] <- apply(smry_v3[ ,-1], 2, function(x) as.numeric(as.character(x)))

###QC
View(smry_v3)
```

```{r}
otm <- cms_tgt2*otp_tgt2/100
s2 <- sapply(1:length(a_number), function(x) {
  solve(1, .78 * (cms_tgt2[x] + target[x]) - otm[x])
})
s2
reqotp <- s2/target
reqotp
```

```{r}
smry_v3 <- replace(smry_v3, smry_v3 < 0, 0)
```
